---
title: "Experiment 2: Main Analysis"
format: html
editor: visual
execute:
  cache: false
---

## Overview

This notebook reproduces all statistical analyses for Experiment 2, organized to match the Results section structure:

1. **Error reduction over time** - Groups improve, chat outperforms slider
2. **Difficulty × condition interactions** - Chat's advantage under epistemic difficulty
3. **Continuous difficulty predictors** - Representativeness as key driver
4. **Message content analysis** - Social information predicts accuracy

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

library(tidyverse)
library(lme4)
library(lmerTest)
conflicts_prefer(lmerTest::lmer)
library(here)
library(emmeans)
library(ggthemes)
library(patchwork)
library(broom.mixed)
library(tidyboot)

theme_set(theme_few())

# Colors for conditions
condition_colors <- c("slider" = "#E69F00", "interactive" = "#009E73")  # Match Exp 1 palette
```

## Load Data

```{r}
#| label: load-data
#| message: false

# Load clean data (produced by 1_data_cleaning.qmd)
d.all <- read_csv(here('analysis/exp2/exp2_clean_data.csv'))

# Derive condition and direction from treatmentName
d.all <- d.all %>%
  mutate(
    condition = ifelse(str_detect(treatmentName.x, "interactive"), "interactive", "slider"),
    direction = ifelse(str_detect(treatmentName.x, "^RS_"), "RS", "SR")
  )

# Shift slider index so round 1 = first round with social information
# In slider condition, round 1 is the initial guess BEFORE seeing others' numbers
# In chat condition, round 1 already includes discussion
# This aligns the conditions for fair comparison

# Create shifted index (baseline will be saved AFTER error is computed below)
d.all <- d.all %>%
  mutate(
    idx_shifted = if_else(condition == "slider", idx - 1, idx)
  )

cat("=== DATA SUMMARY (Intent-to-Treat, with index shift) ===\n")
cat("Participants:", n_distinct(d.all$playerID), "\n")
cat("Games:", n_distinct(d.all$gameID), "\n")
cat("Observations:", nrow(d.all), "\n")

# Check conditions
cat("\nBy condition:\n")
d.all %>%
  distinct(gameID, condition) %>%
  count(condition) %>%
  print()
```

## Compute Actual Network Metrics

We recompute game-level statistics based on the **final state of the network** - only participants who actually responded contribute to the collective information pool.

```{r}
#| label: actual-network-metrics

# Identify active players (those who responded)
active_players <- d.all %>%
  filter(!is.na(guess)) %>%
  distinct(gameID, playerID, nCritters, nRabbits, nSquirrels)

cat("Active players per game:\n")
active_players %>%
  count(gameID, name = "n_active") %>%
  count(n_active) %>%
  print()

# Recompute game-level metrics based on active players
game_metrics_actual <- active_players %>%
  group_by(gameID) %>%
  summarise(
    n_active_players = n(),
    nCrittersGame_actual = sum(nCritters),
    nRabbitsGame_actual = sum(nRabbits),
    mleGame_actual = nRabbitsGame_actual / nCrittersGame_actual,
    .groups = "drop"
  )

# Entropy based on active players
entropy_actual <- active_players %>%
  group_by(gameID) %>%
  mutate(
    total_critters = sum(nCritters),
    p_i = nCritters / total_critters
  ) %>%
  summarise(
    entropy_actual = -sum(p_i * log(p_i + 1e-10)),
    .groups = "drop"
  )

# Representativeness: use the existing column from the data
# This is mean |individual_MLE - collective_MLE|, measuring how misleading individual samples are
# (NOT |mle - 0.5| which measures distance from 50%)
game_metrics_actual <- game_metrics_actual %>%
  left_join(entropy_actual, by = "gameID") %>%
  left_join(
    d.all %>% distinct(gameID, representation) %>% rename(representation_actual = representation),
    by = "gameID"
  )

# Merge back to all participants
d.all <- d.all %>%
  left_join(game_metrics_actual, by = "gameID")

# CRITICAL: Align MLE for direction
# In RS direction games, the majority was squirrels, but participants still report % rabbits
# So the "true" network MLE needs to be inverted for RS games
d.all <- d.all %>%
  mutate(
    mleGame_aligned = ifelse(direction == "RS", 1 - mleGame_actual, mleGame_actual)
  )

cat("\nActual network metrics computed for", n_distinct(game_metrics_actual$gameID), "games\n")
cat("Direction distribution:\n")
d.all %>%
  distinct(gameID, direction) %>%
  count(direction) %>%
  print()
```

## Compute Error

```{r}
#| label: compute-error

# Compute absolute error from ALIGNED network MLE
# This accounts for direction (RS games have inverted MLE)
d.all <- d.all %>%
  mutate(error = abs(guess - mleGame_aligned * 100))

# Now save slider's baseline (pre-social, idx_shifted=0) for visualization
d.slider_baseline <- d.all %>%
  filter(condition == "slider", idx_shifted == 0)

# Filter to only include rounds with social information (idx_shifted >= 1)
d.all <- d.all %>%
  filter(idx_shifted >= 1)

cat("Error summary:\n")
summary(d.all$error)

# Check for magnet values
cat("\nMagnet value frequencies:\n")
d.all %>%
  filter(!is.na(error)) %>%
  mutate(is_magnet = guess %in% c(0, 50, 100)) %>%
  count(condition, is_magnet) %>%
  group_by(condition) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  print()
```

## Create Difficulty Bins (Easy/Hard)

Using the **actual network metrics** for binning:

```{r}
#| label: create-bins

# Median split on each difficulty dimension (using actual network metrics)
d.all <- d.all %>%
  mutate(
    # Median splits for easy/hard classification
    critters_easy = nCrittersGame_actual >= median(nCrittersGame_actual, na.rm = TRUE),
    entropy_easy = entropy_actual >= median(entropy_actual, na.rm = TRUE),
    rep_easy = representation_actual <= median(representation_actual, na.rm = TRUE),  # Lower = easier

    # Create combined difficulty profile
    difficulty_profile = paste0(
      ifelse(critters_easy, "E", "H"),  # Critters
      ifelse(entropy_easy, "E", "H"),   # Entropy
      ifelse(rep_easy, "E", "H")        # Representativeness
    ),

    # Count hard dimensions
    n_hard = (!critters_easy) + (!entropy_easy) + (!rep_easy),

    # Create factor bins for plotting
    critters_bin = factor(ifelse(critters_easy, "Many", "Few"), levels = c("Few", "Many")),
    entropy_bin = factor(ifelse(entropy_easy, "High", "Low"), levels = c("Low", "High")),
    rep_bin = factor(ifelse(rep_easy, "Representative", "Unrepresentative"),
                     levels = c("Representative", "Unrepresentative"))
  )

cat("Difficulty profile distribution:\n")
d.all %>%
  distinct(gameID, difficulty_profile) %>%
  count(difficulty_profile) %>%
  arrange(desc(n)) %>%
  print()

# Add difficulty bins to slider baseline data (for plotting)
d.slider_baseline <- d.slider_baseline %>%
  left_join(
    d.all %>% distinct(gameID, n_hard, critters_bin, entropy_bin, rep_bin),
    by = "gameID"
  )
```

---

# Analysis 1: Error Reduction Over Time

**Key question:** Do groups improve over time? Does chat outperform slider?

## 1.0 Create Effect-Coded Variables

```{r}
#| label: effect-coding

# Effect coding for all analyses
# Condition: -0.5 = chat, +0.5 = slider
# Difficulty: scaled 0-1 and mean-centered (continuous predictors)

# First, verify the difficulty dimensions are continuous (not factorial design)
cat("=== DIFFICULTY DIMENSION DISTRIBUTIONS ===\n")
game_difficulty <- d.all %>%
  distinct(gameID, nCrittersGame_actual, entropy_actual, representation_actual)
cat("Sample size: ", n_distinct(game_difficulty$nCrittersGame_actual), "unique values\n")
cat("Entropy: ", n_distinct(round(game_difficulty$entropy_actual, 3)), "unique values\n")
cat("Representativeness: ", n_distinct(round(game_difficulty$representation_actual, 3)), "unique values\n")
cat("Total unique combinations:", nrow(game_difficulty), "(vs 8 expected for 2×2×2 factorial)\n\n")
cat("Since difficulty dimensions are CONTINUOUS, we use centered continuous predictors.\n\n")

d.all <- d.all %>%
  mutate(
    # Effect-coded condition
    condition_c = ifelse(condition == "slider", 0.5, -0.5),
    idx_c = idx_shifted - mean(idx_shifted, na.rm = TRUE),

    # Scale difficulty to 0-1 range, then mean-center
    # Higher values = MORE DIFFICULT for all three
    critters_scaled = 1 - (nCrittersGame_actual - min(nCrittersGame_actual, na.rm = TRUE)) /
                          (max(nCrittersGame_actual, na.rm = TRUE) - min(nCrittersGame_actual, na.rm = TRUE)),  # Fewer = harder
    entropy_scaled = 1 - (entropy_actual - min(entropy_actual, na.rm = TRUE)) /
                         (max(entropy_actual, na.rm = TRUE) - min(entropy_actual, na.rm = TRUE)),  # Lower entropy = harder
    rep_scaled = (representation_actual - min(representation_actual, na.rm = TRUE)) /
                 (max(representation_actual, na.rm = TRUE) - min(representation_actual, na.rm = TRUE)),  # Higher = harder (less representative)

    # Mean-center
    critters_c = critters_scaled - mean(critters_scaled, na.rm = TRUE),
    entropy_c = entropy_scaled - mean(entropy_scaled, na.rm = TRUE),
    rep_c = rep_scaled - mean(rep_scaled, na.rm = TRUE)
  )

cat("Effect coding applied:\n")
cat("  condition_c: chat = -0.5, slider = +0.5\n")
cat("  idx_c: centered at mean round\n")
cat("  difficulty: scaled 0-1 (higher = harder), then mean-centered\n")
cat("    critters_c range:", round(min(d.all$critters_c), 3), "to", round(max(d.all$critters_c), 3), "\n")
cat("    entropy_c range:", round(min(d.all$entropy_c), 3), "to", round(max(d.all$entropy_c), 3), "\n")
cat("    rep_c range:", round(min(d.all$rep_c), 3), "to", round(max(d.all$rep_c), 3), "\n")
```

## 1.1 Basic Model: Error ~ Round

```{r}
#| label: analysis1-basic

# Linear mixed-effects model: error decreases over time
model1_basic <- lmer(
  error ~ idx_c + (1 + idx_c | gameID/playerID),
  data = d.all
)

cat("=== Model 1: Error ~ Round (centered) ===\n")
summary(model1_basic)
```

## 1.2 Condition Effect: Error ~ Round × Condition (Effect-Coded)

```{r}
#| label: analysis1-condition

# Model with effect-coded condition
model1_condition <- lmer(
  error ~ condition_c * idx_c + (1 + idx_c | gameID/playerID),
  data = d.all
)

cat("=== Model 2: Error ~ Round × Condition (effect-coded) ===\n")
summary(model1_condition)

# Extract key statistics for manuscript
cat("\n=== KEY STATISTICS FOR MANUSCRIPT ===\n")
coefs <- fixef(model1_condition)
se <- sqrt(diag(vcov(model1_condition)))
df <- summary(model1_condition)$coefficients[, "df"]
pvals <- summary(model1_condition)$coefficients[, "Pr(>|t|)"]

cat("Round effect: b =", round(coefs["idx_c"], 2),
    ", SE =", round(se["idx_c"], 2),
    ", p <", ifelse(pvals["idx_c"] < 0.001, ".001", round(pvals["idx_c"], 3)), "\n")
cat("Condition effect: b =", round(coefs["condition_c"], 2),
    ", SE =", round(se["condition_c"], 2),
    ", p =", round(pvals["condition_c"], 4), "\n")
cat("Interaction: b =", round(coefs["condition_c:idx_c"], 2),
    ", SE =", round(se["condition_c:idx_c"], 2),
    ", p =", round(pvals["condition_c:idx_c"], 3), "\n")

# Condition means
cat("\nCondition means:\n")
d.all %>%
  group_by(condition) %>%
  summarise(mean_error = round(mean(error, na.rm = TRUE), 1),
            sd_error = round(sd(error, na.rm = TRUE), 1),
            n = n()) %>%
  print()

# Final round means
cat("\nFinal round means:\n")
d.all %>%
  group_by(condition) %>%
  filter(idx_shifted == max(idx_shifted, na.rm = TRUE)) %>%
  summarise(final_error = round(mean(error, na.rm = TRUE), 1),
            n = n()) %>%
  print()
```

## 1.3 Figure 5A: Error Over Time by Condition

```{r}
#| label: fig5a
#| fig-width: 8
#| fig-height: 5

# Compute bootstrapped means
error_by_round <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(condition, idx_shifted) %>%
  tidyboot_mean(error, na.rm = TRUE)

fig5a <- ggplot(error_by_round, aes(x = idx_shifted, y = empirical_stat, color = condition)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper, fill = condition),
              alpha = 0.2, color = NA) +
  geom_point(size = 2) +
  scale_color_manual(values = condition_colors,
                     labels = c("slider" = "Slider", "interactive" = "Chat")) +
  scale_fill_manual(values = condition_colors,
                    labels = c("slider" = "Slider", "interactive" = "Chat")) +
  labs(
    x = "Round",
    y = "Mean Absolute Error",
    color = "Condition",
    fill = "Condition",
    title = "A. Error decreases over time"
  ) +
  theme(legend.position = "bottom")

fig5a
```

## 1.4 Figure 5B: Final Round Error Distribution

```{r}
#| label: fig5b
#| fig-width: 6
#| fig-height: 5

# Final round only (each condition's max shifted index)
d.final <- d.all %>%
  group_by(condition) %>%
  filter(idx_shifted == max(idx_shifted, na.rm = TRUE)) %>%
  ungroup()

fig5b <- ggplot(d.final, aes(x = condition, y = error, fill = condition)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.2, alpha = 0.8, outlier.shape = NA) +
  scale_fill_manual(values = condition_colors,
                    labels = c("slider" = "Slider", "interactive" = "Chat")) +
  scale_x_discrete(labels = c("slider" = "Slider", "interactive" = "Chat")) +
  labs(
    x = "Condition",
    y = "Final Round Error",
    title = "B. Final round error distribution"
  ) +
  theme(legend.position = "none")

fig5b
```

---

# Analysis 2: Difficulty × Condition Interactions (Binned)

**Key question:** Does chat's advantage emerge specifically under epistemic difficulty?

## 2.1 Three-Way Hard Condition

```{r}
#| label: analysis2-hard-all

# Filter to all-hard (HHH) and all-easy (EEE) for clearest comparison
d.extreme <- d.all %>%
  filter(difficulty_profile %in% c("HHH", "EEE"))

cat("Sample sizes for extreme conditions:\n")
d.extreme %>%
  distinct(gameID, condition, difficulty_profile) %>%
  count(difficulty_profile, condition) %>%
  print()

# Model: Error ~ Condition × Difficulty Profile
model2_extreme <- lmer(
  error ~ condition * difficulty_profile + (1 | gameID/playerID),
  data = d.extreme
)

cat("\n=== Model: Extreme Difficulty Profiles ===\n")
summary(model2_extreme)
```

## 2.2 Full Difficulty Profile Analysis

```{r}
#| label: analysis2-full

# All 8 difficulty profiles
model2_full <- lmer(
  error ~ condition * difficulty_profile + (1 | gameID/playerID),
  data = d.all
)

cat("=== Model: All Difficulty Profiles ===\n")
summary(model2_full)

# Estimated marginal means
cat("\n=== Estimated Marginal Means ===\n")
emmeans(model2_full, ~ condition | difficulty_profile)
```

## 2.3 Two-Way Difficulty Combinations

```{r}
#| label: analysis2-twoway

# Create two-way difficulty indicators
d.all <- d.all %>%
  mutate(
    hard_entropy_rep = (!entropy_easy) & (!rep_easy),
    hard_critters_rep = (!critters_easy) & (!rep_easy),
    hard_entropy_critters = (!entropy_easy) & (!critters_easy)
  )

# Hard entropy + hard representativeness
cat("=== Hard Entropy × Hard Representativeness ===\n")
d.subset <- d.all %>% filter(hard_entropy_rep)
if(nrow(d.subset) > 100) {
  model_ent_rep <- lmer(error ~ condition + (1 | gameID/playerID), data = d.subset)
  print(summary(model_ent_rep))
}

# Hard critters + hard representativeness
cat("\n=== Hard Critters × Hard Representativeness ===\n")
d.subset <- d.all %>% filter(hard_critters_rep)
if(nrow(d.subset) > 100) {
  model_crit_rep <- lmer(error ~ condition + (1 | gameID/playerID), data = d.subset)
  print(summary(model_crit_rep))
}

# Hard entropy + hard critters (easy representativeness)
cat("\n=== Hard Entropy × Hard Critters (Easy Rep) ===\n")
d.subset <- d.all %>% filter(hard_entropy_critters & rep_easy)
if(nrow(d.subset) > 100) {
  model_ent_crit <- lmer(error ~ condition + (1 | gameID/playerID), data = d.subset)
  print(summary(model_ent_crit))
}
```

## 2.4 Figure 6A: Difficulty Profile × Condition

```{r}
#| label: fig6a
#| fig-width: 10
#| fig-height: 6

# Compute mean error by difficulty profile and condition (bootstrapped CIs)
difficulty_summary <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(difficulty_profile, condition) %>%
  tidyboot_mean(error, na.rm = TRUE) %>%
  # Order by number of hard dimensions
  mutate(
    n_hard = str_count(difficulty_profile, "H"),
    difficulty_profile = fct_reorder(difficulty_profile, n_hard)
  )

fig6a <- ggplot(difficulty_summary,
                aes(x = difficulty_profile, y = empirical_stat, fill = condition)) +
  geom_col(position = position_dodge(0.8), width = 0.7) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(0.8), width = 0.2) +
  scale_fill_manual(values = condition_colors,
                    labels = c("slider" = "Slider", "interactive" = "Chat")) +
  labs(
    x = "Difficulty Profile (Critters/Entropy/Representativeness)",
    y = "Mean Absolute Error",
    fill = "Condition",
    title = "A. Error by difficulty profile and condition"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

fig6a
```

---

# Analysis 3: Difficulty × Condition Interactions (Centered Continuous)

**Key question:** Which difficulty dimension most strongly moderates chat's advantage?

**Modeling approach:** We use centered continuous predictors for difficulty (scaled 0-1, then mean-centered) so that:
- Main effects are interpretable at the average level of all other predictors
- Coefficients represent the change in error per unit change in difficulty (full 0-1 range)
- Interaction coefficients show how much the condition effect changes across the full range of difficulty

## 3.0 Correlation Between Difficulty Dimensions

```{r}
#| label: difficulty-correlations

# Correlation between entropy and representativeness (reported in manuscript)
game_level <- d.all %>%
  distinct(gameID, entropy_actual, representation_actual)

r_entropy_rep <- cor(game_level$entropy_actual, game_level$representation_actual,
                     use = "complete.obs")
cat("Correlation between entropy and representativeness: r =", round(r_entropy_rep, 2), "\n")
```

## 3.1 Full Factorial Model

Testing all main effects and interactions between condition and difficulty dimensions:

```{r}
#| label: analysis3-full-factorial

# Full factorial model with effect-coded predictors
# Include round index to control for learning over time
# All two-way interactions between condition and difficulty dimensions
# Plus difficulty × difficulty interactions
model3_full <- lmer(
  error ~ idx_c +
          condition_c * critters_c +
          condition_c * entropy_c +
          condition_c * rep_c +
          critters_c * entropy_c +
          critters_c * rep_c +
          entropy_c * rep_c +
    (1 | gameID/playerID),
  data = d.all
)

cat("=== Full Factorial Model (Effect-Coded) ===\n")
cat("Coding: -0.5 = easy/chat, +0.5 = hard/slider\n\n")
summary(model3_full)
```

## 3.2 Extract Key Statistics

```{r}
#| label: analysis3-key-stats

coefs <- summary(model3_full)$coefficients

# Helper function to format statistics
format_stat <- function(coefs, term) {
  b <- coefs[term, "Estimate"]
  t_val <- coefs[term, "t value"]
  df <- coefs[term, "df"]
  p <- coefs[term, "Pr(>|t|)"]
  sig <- ifelse(p < 0.001, "***", ifelse(p < 0.01, "**", ifelse(p < 0.05, "*", "")))
  p_str <- ifelse(p < 0.001, "< .001", sprintf("= %.3f", p))
  sprintf("b = %5.2f, t(%3.0f) = %5.2f, p %s %s", b, df, t_val, p_str, sig)
}

cat("=== MAIN EFFECTS ===\n")
cat("(Averaging across all other factors)\n\n")
cat("  Round (centered):   ", format_stat(coefs, "idx_c"), "\n")
cat("  Condition:          ", format_stat(coefs, "condition_c"), "\n")
cat("  Sample size:        ", format_stat(coefs, "critters_c"), "\n")
cat("  Entropy:            ", format_stat(coefs, "entropy_c"), "\n")
cat("  Representativeness: ", format_stat(coefs, "rep_c"), "\n")

cat("\n=== CONDITION × DIFFICULTY INTERACTIONS ===\n")
cat("(Does slider's disadvantage change with difficulty?)\n\n")
cat("  Cond × Sample size: ", format_stat(coefs, "condition_c:critters_c"), "\n")
cat("  Cond × Entropy:     ", format_stat(coefs, "condition_c:entropy_c"), "\n")
cat("  Cond × Rep:         ", format_stat(coefs, "condition_c:rep_c"), "\n")

cat("\n=== DIFFICULTY × DIFFICULTY INTERACTIONS ===\n")
cat("(Do sources of difficulty compound?)\n\n")
cat("  Sample × Entropy:   ", format_stat(coefs, "critters_c:entropy_c"), "\n")
cat("  Sample × Rep:       ", format_stat(coefs, "critters_c:rep_c"), "\n")
cat("  Entropy × Rep:      ", format_stat(coefs, "entropy_c:rep_c"), "\n")

# Store for later use
model3_interactions <- model3_full
```

## 3.3 Cell Means for Representativeness × Condition

```{r}
#| label: analysis3-cell-means

cat("=== CELL MEANS: Representativeness × Condition ===\n\n")
d.all %>%
  filter(!is.na(error)) %>%
  group_by(rep_bin, condition) %>%
  summarise(
    mean_error = round(mean(error, na.rm = TRUE), 1),
    sd_error = round(sd(error, na.rm = TRUE), 1),
    n = n(),
    .groups = "drop"
  ) %>%
  print()
```

## 3.4 Create Coefficient Table for Manuscript

```{r}
#| label: analysis3-coef-table

# Extract interaction coefficients for figure
coef_table <- tidy(model3_full, effects = "fixed", conf.int = TRUE) %>%
  filter(str_detect(term, "condition_c:") & !str_detect(term, ":.*:")) %>%
  mutate(
    dimension = case_when(
      str_detect(term, "critters") ~ "Sample Size",
      str_detect(term, "entropy") ~ "Distribution",
      str_detect(term, "rep") ~ "Representativeness"
    )
  )

cat("Interaction coefficients for figure:\n")
print(coef_table %>% select(dimension, estimate, conf.low, conf.high, p.value))
```

## 3.5 Figure 6B: Interaction Coefficients

```{r}
#| label: fig6b
#| fig-width: 6
#| fig-height: 4

# Coefficient plot
fig6b <- ggplot(coef_table, aes(x = reorder(dimension, abs(estimate)),
                                 y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    x = "",
    y = "Condition × Dimension Interaction (β)",
    title = "B. Interaction coefficients"
  )

fig6b
```

---

# Analysis 4: Message Content

**Key question:** What aspects of chat communication predict accuracy?

**Coding procedure:** Two coders independently categorized each message along two dimensions:

- **Source**: Personal (own observation) vs. Social (relaying information from others)
- **Type**: Numeric (specific counts) vs. Belief (estimates/opinions)

This yields four main categories:
- `pn` = personal numeric (e.g., "I saw 3 rabbits and 2 squirrels")
- `pb` = personal belief (e.g., "I think it's around 70%")
- `sn` = social numeric (e.g., "My partner said they saw 5 rabbits")
- `sb` = social belief (e.g., "They thought it was 60%")

Messages could receive compound codes if they contained multiple types (e.g., "pnsb" = personal numeric + social belief).

## 4.1 Load Coded Messages

```{r}
#| label: load-messages
#| message: false

# Load coded message data
d.messages <- read_csv(here('analysis/exp2/message_data/coded.csv'))

cat("=== MESSAGE DATA ===\n")
cat("Total messages:", nrow(d.messages), "\n")
cat("Unique games:", n_distinct(d.messages$gameID), "\n")

# Coding categories (including compound codes)
cat("\nCoding categories:\n")
d.messages %>%
  count(coding) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  arrange(desc(n)) %>%
  print(n = 15)
```

## 4.2 Aggregate Message Counts per Game

```{r}
#| label: message-counts

# Count message types per game
# Use str_detect to capture compound codes (e.g., "pnsb" contains both pn and sb)
message_counts <- d.messages %>%
  group_by(gameID) %>%
  summarise(
    total_messages = n(),
    pn_count = sum(str_detect(coding, "pn"), na.rm = TRUE),  # personal numeric
    pb_count = sum(str_detect(coding, "pb"), na.rm = TRUE),  # personal belief
    sn_count = sum(str_detect(coding, "sn"), na.rm = TRUE),  # social numeric
    sb_count = sum(str_detect(coding, "sb"), na.rm = TRUE),  # social belief
    o_count = sum(coding == "o", na.rm = TRUE),              # other (exact match)
    q_count = sum(str_detect(coding, "q"), na.rm = TRUE),    # question
    .groups = "drop"
  ) %>%
  mutate(
    # Proportions of messages containing each type
    p_social = (sn_count + sb_count) / total_messages,
    p_personal = (pn_count + pb_count) / total_messages
  )

cat("Message count summary:\n")
summary(message_counts %>% select(total_messages, pn_count:q_count))

cat("\nProportion social information:\n")
cat("Mean:", round(mean(message_counts$p_social), 3), "\n")
cat("Range:", round(range(message_counts$p_social), 3), "\n")
```

## 4.3 Merge with Game-Level Error

```{r}
#| label: merge-messages

# Compute final-round error per game (using shifted index)
game_error <- d.all %>%
  group_by(gameID, condition) %>%
  filter(idx_shifted == max(idx_shifted, na.rm = TRUE)) %>%
  group_by(gameID) %>%
  summarise(
    avg_error = mean(error, na.rm = TRUE),
    .groups = "drop"
  )

# Merge
d.message_analysis <- message_counts %>%
  inner_join(game_error, by = "gameID")

cat("Games with both message and error data:", nrow(d.message_analysis), "\n")
```

## 4.4 Social Information Predicts Accuracy

**Model specification:** We predict log-transformed final error from the proportion of messages containing social information, controlling for personal information sharing. Log transformation accounts for the right-skewed error distribution.

```{r}
#| label: analysis4-social

# Primary model: proportion of social information predicts accuracy
# Outcome: log(avg_error + 1) to handle skew
# Predictors: p_social (proportion social), p_personal (proportion personal)
model4_social <- lm(
  log(avg_error + 1) ~ p_social + p_personal,
  data = d.message_analysis
)

cat("=== Primary Model: Proportion Social Information → log(Error) ===\n")
summary(model4_social)

cat("\n=== Key Result for Manuscript ===\n")
coef_social <- coef(summary(model4_social))["p_social", ]
cat("p_social coefficient: b =", round(coef_social["Estimate"], 2),
    ", p =", format.pval(coef_social["Pr(>|t|)"], digits = 3), "\n")
cat("Interpretation: Higher proportion of social information → lower log(error)\n")

# Secondary model: raw counts by message type
model4_types <- lm(
  log(avg_error + 1) ~ pn_count + pb_count + sn_count + sb_count,
  data = d.message_analysis
)

cat("\n=== Secondary Model: Message Counts → log(Error) ===\n")
summary(model4_types)
```

## 4.5 Figure 7B: Message Type Frequencies

```{r}
#| label: fig7b
#| fig-width: 6
#| fig-height: 4

# Frequency plot - filter to main coding categories only
freq_data <- d.messages %>%
  filter(coding %in% c("pn", "pb", "sn", "sb", "o", "q")) %>%
  count(coding) %>%
  mutate(
    label = case_when(
      coding == "pn" ~ "Personal\nNumeric",
      coding == "pb" ~ "Personal\nBelief",
      coding == "sn" ~ "Social\nNumeric",
      coding == "sb" ~ "Social\nBelief",
      coding == "o" ~ "Other",
      coding == "q" ~ "Question",
      TRUE ~ coding
    ),
    type = ifelse(coding %in% c("sn", "sb"), "Social", "Personal/Other")
  )

fig7b <- ggplot(freq_data, aes(x = reorder(label, -n), y = n, fill = type)) +
  geom_col() +
  scale_fill_manual(values = c("Social" = "#0072B2", "Personal/Other" = "gray60")) +
  labs(
    x = "",
    y = "Count",
    fill = "",
    title = "B. Message type frequencies"
  ) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 9))

fig7b
```

## 4.6 Figure 7C: Social Information vs Error

```{r}
#| label: fig7c
#| fig-width: 6
#| fig-height: 5

fig7c <- ggplot(d.message_analysis, aes(x = p_social, y = avg_error)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "#0072B2") +
  labs(
    x = "Proportion Social Information (sn + sb)",
    y = "Final Round Error",
    title = "C. Social information predicts accuracy"
  )

fig7c
```

---

# Combine and Save Figures

## Figure 5: Combined Results Figure (Time + Difficulty)

This is the main "money plot" combining error over time with the three difficulty dimensions.

```{r}
#| label: save-fig5
#| cache: false
#| fig-width: 12
#| fig-height: 3.5

# Panel A: Error over time COLLAPSED (bootstrapped CIs)
error_by_round <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(condition, idx_shifted) %>%
  tidyboot_mean(error, na.rm = TRUE)

# Compute slider baseline (pre-social, collapsed)
slider_baseline_collapsed <- d.slider_baseline %>%
  filter(!is.na(error)) %>%
  tidyboot_mean(error, na.rm = TRUE) %>%
  mutate(condition = "slider", idx_shifted = 0)

# Get first points for each condition for dashed lines
first_points <- error_by_round %>%
  filter(idx_shifted == 1) %>%
  select(condition, empirical_stat)

# Create baseline segments for BOTH conditions (connecting baseline to round 1)
baseline_segments <- first_points %>%
  mutate(
    baseline_y = slider_baseline_collapsed$empirical_stat,
    x_start = 0,
    x_end = 1
  )

fig5_a <- ggplot(error_by_round,
                 aes(x = idx_shifted, y = empirical_stat, color = condition)) +
  # Add dashed lines from baseline to first social-info round (both conditions)
  geom_segment(data = baseline_segments,
               aes(x = x_start, xend = x_end, y = baseline_y, yend = empirical_stat, color = condition),
               linetype = "dashed", linewidth = 0.8) +
  # Add baseline point (slider only)
  geom_point(data = slider_baseline_collapsed,
             aes(x = idx_shifted, y = empirical_stat),
             size = 2, color = condition_colors["slider"]) +
  # Main plot elements - connect points with lines (no smoothing)
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper, fill = condition),
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = condition_colors,
                     labels = c("interactive" = "Chat", "slider" = "Slider"),
                     name = NULL) +
  scale_fill_manual(values = condition_colors,
                    labels = c("interactive" = "Chat", "slider" = "Slider"),
                    name = NULL) +
  labs(x = "Round", y = "Mean Absolute Error", tag = "A") +
  theme_few() +
  theme(legend.position = c(0.75, 0.85),
        legend.background = element_blank(),
        legend.key.size = unit(0.4, "cm"),
        aspect.ratio = 1)

fig5_a

# Save Panel A separately
ggsave(here("figures/exp2_fig5a.pdf"), fig5_a, width = 4, height = 4)

# Panel B: All difficulty dimensions as facets (MARGINAL MEANS from emmeans)
# Use emmeans to control for confounds - each dimension controls for the other two

# Sample Size: control for entropy and representativeness
model_size <- lmer(error ~ condition * critters_bin + entropy_c + rep_c + (1 | gameID),
                   data = d.all)
emm_size <- emmeans(model_size, ~ condition * critters_bin)
summary_critters <- as.data.frame(confint(emm_size)) %>%
  mutate(dimension = "Total Sample Size",
         level = as.character(critters_bin)) %>%
  rename(empirical_stat = emmean, ci_lower = asymp.LCL, ci_upper = asymp.UCL)

# Distribution (entropy): control for sample size and representativeness
model_entropy <- lmer(error ~ condition * entropy_bin + critters_c + rep_c + (1 | gameID),
                      data = d.all)
emm_entropy <- emmeans(model_entropy, ~ condition * entropy_bin)
summary_entropy <- as.data.frame(confint(emm_entropy)) %>%
  mutate(dimension = "Distribution",
         level = as.character(entropy_bin)) %>%
  rename(empirical_stat = emmean, ci_lower = asymp.LCL, ci_upper = asymp.UCL)

# Representativeness: control for sample size and entropy
model_rep <- lmer(error ~ condition * rep_bin + critters_c + entropy_c + (1 | gameID),
                  data = d.all)
emm_rep <- emmeans(model_rep, ~ condition * rep_bin)
summary_rep <- as.data.frame(confint(emm_rep)) %>%
  mutate(dimension = "Representativeness",
         level = as.character(rep_bin)) %>%
  rename(empirical_stat = emmean, ci_lower = asymp.LCL, ci_upper = asymp.UCL)

# Combine all dimensions
summary_all <- bind_rows(summary_critters, summary_entropy, summary_rep) %>%
  mutate(
    dimension = factor(dimension, levels = c("Total Sample Size", "Distribution", "Representativeness")),
    # Rename Low/High to Uneven/Even for distribution
    level = recode(level, "Low" = "Uneven", "High" = "Even"),
    # Order all as [hard, easy]: Few→Many, Uneven→Even, Unrepresentative→Representative
    level = factor(level, levels = c("Few", "Many", "Uneven", "Even", "Unrepresentative", "Representative"))
  )

fig5_b <- ggplot(summary_all, aes(x = level, y = empirical_stat, color = condition, group = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.15) +
  facet_wrap(~ dimension, scales = "free_x", nrow = 1) +
  scale_color_manual(values = condition_colors,
                     labels = c("interactive" = "Chat", "slider" = "Slider"),
                     name = NULL) +
  labs(x = NULL, y = "Mean Absolute Error", tag = "B") +
  theme_few() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 9, angle = 30, hjust = 1),
        strip.text = element_text(size = 10)) +
  coord_cartesian(ylim = c(15, 34))

# Combine Panel A and B (legend stays in Panel A corner only)
fig5 <- (fig5_a | fig5_b) +
  plot_layout(widths = c(1, 1.5))

fig5

ggsave(here("figures/fig5.pdf"), fig5, width = 10, height = 3.5)
ggsave(here("figures/fig5.png"), fig5, width = 10, height = 3.5, dpi = 300)
cat("Saved Figure 5 (combined results)\n")
```

## Figure 6: Message Content

```{r}
#| label: save-fig6
#| cache: false
#| fig-width: 9
#| fig-height: 4

# Natural breaks: 0%, 1-10%, >10%
d.message_analysis <- d.message_analysis %>%
  mutate(
    social_level = factor(
      case_when(
        p_social == 0 ~ "0%",
        p_social <= 0.10 ~ "1–10%",
        p_social > 0.10 ~ ">10%"
      ),
      levels = c("0%", "1–10%", ">10%")
    )
  )

# Panel A: Error by social relay level (bootstrapped CIs)
social_summary <- d.message_analysis %>%
  filter(!is.na(social_level)) %>%
  group_by(social_level) %>%
  tidyboot_mean(avg_error, na.rm = TRUE)

fig6_a <- ggplot(social_summary, aes(x = social_level, y = empirical_stat)) +
  geom_col(fill = "#0072B2", alpha = 0.7, width = 0.6) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  labs(x = "% Messages Relaying Social Information", y = "Final Round Error", tag = "A") +
  theme_few() +
  theme(axis.text.x = element_text(size = 12)) +
  coord_cartesian(ylim = c(0, 25))

# Panel B: Example messages (as a table/text panel)
# Create a simple visualization showing example messages
examples <- tibble(
  Category = c("Personal\nNumeric", "Personal\nNumeric",
               "Social\nNumeric", "Social\nNumeric"),
  Example = c(
    "\"4 rabbits, 2 squirrels\"",
    "\"I saw more rabbits\"",
    "\"My neighbor saw 11 rabbits\nand 1 squirrel\"",
    "\"I have gathered 22 rabbits\nfrom everyone\""
  ),
  Type = c("Personal", "Personal", "Social", "Social")
)

fig6_b <- ggplot(examples, aes(x = 1, y = rev(seq_along(Example)))) +

  geom_label(aes(label = Example, fill = Type),
             hjust = 0, size = 3, label.padding = unit(0.4, "lines")) +
  scale_fill_manual(values = c("Personal" = "gray85", "Social" = "#0072B2"),
                    guide = "none") +
  scale_x_continuous(limits = c(0.9, 3)) +
  labs(tag = "B") +
  theme_void() +
  theme(plot.tag = element_text(size = 14, face = "bold"))

fig6 <- fig6_a + fig6_b + plot_layout(widths = c(1.2, 1))

fig6

ggsave(here("figures/fig6.pdf"), fig6, width = 9, height = 4)
ggsave(here("figures/fig6.png"), fig6, width = 9, height = 4, dpi = 300)
cat("Saved Figure 6 (message content)\n")
```

---

# Summary Statistics for Manuscript

```{r}
#| label: summary-stats

cat("\n========================================\n")
cat("SUMMARY STATISTICS FOR MANUSCRIPT\n")
cat("========================================\n\n")

cat("=== Sample ===\n")
cat("N participants:", n_distinct(d.all$playerID), "\n")
cat("N games:", n_distinct(d.all$gameID), "\n")

cat("\n=== Analysis 1: Error over time ===\n")
cat("Round effect: b =", round(fixef(model1_basic)["idx"], 2), "\n")

cat("\n=== Analysis 2: Condition effect ===\n")
# Extract from model1_condition

cat("\n=== Analysis 3: Interactions ===\n")
print(coef_table %>% select(dimension, estimate, conf.low, conf.high, p.value))

cat("\n=== Analysis 4: Message content ===\n")
cat("Social information effect from model4_social\n")
```

---

# Robustness Analyses

## Robustness 1: Exclude Low Response Rate Participants

Following Experiment 1's pre-registered criterion, we exclude participants with <80% response rate.

```{r}
#| label: robustness-response-rate

# Calculate response rates per participant
response_rates <- d.all %>%
  bind_rows(d.slider_baseline) %>%  # Include baseline for full picture
  group_by(playerID) %>%
  summarise(
    total_possible = max(idx_shifted, na.rm = TRUE) + 1,
    responses = sum(!is.na(guess)),
    response_rate = responses / 8,  # 8 rounds total in Exp2
    .groups = "drop"
  )

cat("Response rate distribution:\n")
print(summary(response_rates$response_rate))

# Apply 80% threshold
exclude_low_response <- response_rates %>%
  filter(response_rate < 0.80) %>%
  pull(playerID)

cat("\nParticipants excluded (response rate < 80%):", length(exclude_low_response), "\n")

# Create filtered dataset
d.robust1 <- d.all %>%
  filter(!(playerID %in% exclude_low_response))

cat("Remaining after exclusion:", n_distinct(d.robust1$playerID), "participants,",
    n_distinct(d.robust1$gameID), "games\n")

# Re-run key model with effect-coded variables
if(nrow(d.robust1) > 100) {
  model_robust1 <- lmer(
    error ~ condition_c * rep_c + (1 | gameID/playerID),
    data = d.robust1
  )

  cat("\n=== Robustness: Condition × Representativeness (80% response rate) ===\n")
  coef_robust1 <- fixef(model_robust1)
  cat("Condition × Rep interaction: b =",
      round(coef_robust1["condition_c:rep_c"], 2), "\n")
  print(summary(model_robust1)$coefficients)
}
```

## Robustness 2: Exclude Magnet Values

Excluding guesses of exactly 0%, 50%, or 100% which may indicate inattention.

```{r}
#| label: robustness-magnet

# Count magnet values
cat("Magnet value frequencies:\n")
d.all %>%
  mutate(magnet_type = case_when(
    guess == 0 ~ "0%",
    guess == 50 ~ "50%",
    guess == 100 ~ "100%",
    TRUE ~ "Other"
  )) %>%
  count(magnet_type) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  print()

# Create filtered dataset
d.robust2 <- d.all %>%
  filter(!guess %in% c(0, 50, 100))

cat("\nAfter excluding magnet values:", nrow(d.robust2), "observations\n")

# Re-run key model with effect-coded variables
if(nrow(d.robust2) > 100) {
  model_robust2 <- lmer(
    error ~ condition_c * rep_c + (1 | gameID/playerID),
    data = d.robust2
  )

  cat("\n=== Robustness: Condition × Representativeness (no magnet values) ===\n")
  coef_robust2 <- fixef(model_robust2)
  cat("Condition × Rep interaction: b =",
      round(coef_robust2["condition_c:rep_c"], 2), "\n")
  print(summary(model_robust2)$coefficients)
}
```

## Robustness 3: Full Games Only (4 Active Players)

Restricting to games where all 4 players remained active.

```{r}
#| label: robustness-full-games

# Count active players per game
games_by_size <- d.all %>%
  group_by(gameID) %>%
  summarise(n_active = n_distinct(playerID), .groups = "drop")

cat("Games by number of active players:\n")
print(table(games_by_size$n_active))

# Filter to full games only
full_games <- games_by_size %>%
  filter(n_active == 4) %>%
  pull(gameID)

d.robust3 <- d.all %>%
  filter(gameID %in% full_games)

cat("\nFull games (4 players):", length(full_games), "games,",
    nrow(d.robust3), "observations\n")

# Re-run key model with effect-coded variables
if(nrow(d.robust3) > 100) {
  model_robust3 <- lmer(
    error ~ condition_c * rep_c + (1 | gameID/playerID),
    data = d.robust3
  )

  cat("\n=== Robustness: Condition × Representativeness (full games only) ===\n")
  coef_robust3 <- fixef(model_robust3)
  cat("Condition × Rep interaction: b =",
      round(coef_robust3["condition_c:rep_c"], 2), "\n")
  print(summary(model_robust3)$coefficients)
}
```

## Robustness Summary

```{r}
#| label: robustness-summary

cat("\n========================================\n")
cat("ROBUSTNESS SUMMARY: Condition × Representativeness Interaction\n")
cat("========================================\n\n")

# Helper to extract full statistics
get_stats <- function(model, term) {
  if(is.null(model)) return(list(b=NA, t=NA, df=NA, p=NA))
  coefs <- summary(model)$coefficients
  if(term %in% rownames(coefs)) {
    list(
      b = coefs[term, "Estimate"],
      t = coefs[term, "t value"],
      df = coefs[term, "df"],
      p = coefs[term, "Pr(>|t|)"]
    )
  } else {
    list(b=NA, t=NA, df=NA, p=NA)
  }
}

# Extract from all models
models <- list(
  "Main (ITT)" = model3_interactions,
  "80% Response Rate" = if(exists("model_robust1")) model_robust1 else NULL,
  "No Magnet Values" = if(exists("model_robust2")) model_robust2 else NULL,
  "Full Games Only" = if(exists("model_robust3")) model_robust3 else NULL
)

results <- map_dfr(names(models), function(name) {
  stats <- get_stats(models[[name]], "condition_c:rep_c")
  tibble(
    Analysis = name,
    b = stats$b,
    t = stats$t,
    df = stats$df,
    p = stats$p
  )
})

# Format for display
results <- results %>%
  mutate(
    `t(df)` = sprintf("t(%.0f) = %.2f", df, t),
    p_formatted = ifelse(p < 0.001, "< .001", sprintf("= %.3f", p)),
    Significant = ifelse(p < 0.05, "*", "")
  )

cat("Condition × Representativeness interaction:\n\n")
results %>%
  select(Analysis, b, `t(df)`, p_formatted, Significant) %>%
  mutate(b = round(b, 2)) %>%
  print()
```
