---
title: "Experiment 2: Main Analysis"
format: html
editor: visual
---

## Overview

This notebook reproduces all statistical analyses for Experiment 2, organized to match the Results section structure:

1. **Error reduction over time** - Groups improve, chat outperforms slider
2. **Difficulty × condition interactions** - Chat's advantage under epistemic difficulty
3. **Continuous difficulty predictors** - Representativeness as key driver
4. **Message content analysis** - Social information predicts accuracy

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

library(tidyverse)
library(lme4)
library(lmerTest)
conflicts_prefer(lmerTest::lmer)
library(here)
library(emmeans)
library(ggthemes)
library(patchwork)
library(broom.mixed)
library(tidyboot)

theme_set(theme_few())

# Colors for conditions
condition_colors <- c("slider" = "#E69F00", "interactive" = "#0072B2")
```

## Load Data

```{r}
#| label: load-data
#| message: false

# Load full data (intent-to-treat, no exclusions)
d.all <- read_csv(here('analysis/exp2/exp2_clean_data_jan17.csv'))

# Create condition variable
d.all <- d.all %>%
  mutate(condition = ifelse(str_detect(treatmentName.x, "interactive"), "interactive", "slider"))

cat("=== DATA SUMMARY (Intent-to-Treat) ===\n")
cat("Participants:", n_distinct(d.all$playerID), "\n")
cat("Games:", n_distinct(d.all$gameID), "\n")
cat("Observations:", nrow(d.all), "\n")

# Check conditions
cat("\nBy condition:\n")
d.all %>%
  distinct(gameID, condition) %>%
  count(condition) %>%
  print()
```

## Compute Actual Network Metrics

We recompute game-level statistics based on the **final state of the network** - only participants who actually responded contribute to the collective information pool.

```{r}
#| label: actual-network-metrics

# Identify active players (those who responded)
active_players <- d.all %>%
  filter(!is.na(guess)) %>%
  distinct(gameID, playerID, nCritters, nRabbits, nSquirrels)

cat("Active players per game:\n")
active_players %>%
  count(gameID, name = "n_active") %>%
  count(n_active) %>%
  print()

# Recompute game-level metrics based on active players
game_metrics_actual <- active_players %>%
  group_by(gameID) %>%
  summarise(
    n_active_players = n(),
    nCrittersGame_actual = sum(nCritters),
    nRabbitsGame_actual = sum(nRabbits),
    mleGame_actual = nRabbitsGame_actual / nCrittersGame_actual,
    .groups = "drop"
  )

# Entropy based on active players
entropy_actual <- active_players %>%
  group_by(gameID) %>%
  mutate(
    total_critters = sum(nCritters),
    p_i = nCritters / total_critters
  ) %>%
  summarise(
    entropy_actual = -sum(p_i * log(p_i + 1e-10)),
    .groups = "drop"
  )

# Representativeness: deviation from 50%
game_metrics_actual <- game_metrics_actual %>%
  left_join(entropy_actual, by = "gameID") %>%
  mutate(representation_actual = abs(mleGame_actual - 0.5))

# Merge back to all participants
d.all <- d.all %>%
  left_join(game_metrics_actual, by = "gameID")

cat("\nActual network metrics computed for", n_distinct(game_metrics_actual$gameID), "games\n")
```

## Compute Error

```{r}
#| label: compute-error

# Compute absolute error from ACTUAL network MLE
d.all <- d.all %>%
  mutate(error = abs(guess - mleGame_actual * 100))

cat("Error summary:\n")
summary(d.all$error)

# Check for magnet values
cat("\nMagnet value frequencies:\n")
d.all %>%
  filter(!is.na(error)) %>%
  mutate(is_magnet = guess %in% c(0, 50, 100)) %>%
  count(condition, is_magnet) %>%
  group_by(condition) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  print()
```

## Create Difficulty Bins (Easy/Hard)

Using the **actual network metrics** for binning:

```{r}
#| label: create-bins

# Median split on each difficulty dimension (using actual network metrics)
d.all <- d.all %>%
  mutate(
    # Median splits for easy/hard classification
    critters_easy = nCrittersGame_actual >= median(nCrittersGame_actual, na.rm = TRUE),
    entropy_easy = entropy_actual >= median(entropy_actual, na.rm = TRUE),
    rep_easy = representation_actual <= median(representation_actual, na.rm = TRUE),  # Lower = easier

    # Create combined difficulty profile
    difficulty_profile = paste0(
      ifelse(critters_easy, "E", "H"),  # Critters
      ifelse(entropy_easy, "E", "H"),   # Entropy
      ifelse(rep_easy, "E", "H")        # Representativeness
    ),

    # Count hard dimensions
    n_hard = (!critters_easy) + (!entropy_easy) + (!rep_easy),

    # Create factor bins for plotting
    critters_bin = factor(ifelse(critters_easy, "Many", "Few"), levels = c("Few", "Many")),
    entropy_bin = factor(ifelse(entropy_easy, "High", "Low"), levels = c("Low", "High")),
    rep_bin = factor(ifelse(rep_easy, "Representative", "Unrepresentative"),
                     levels = c("Representative", "Unrepresentative"))
  )

cat("Difficulty profile distribution:\n")
d.all %>%
  distinct(gameID, difficulty_profile) %>%
  count(difficulty_profile) %>%
  arrange(desc(n)) %>%
  print()
```

---

# Analysis 1: Error Reduction Over Time

**Key question:** Do groups improve over time? Does chat outperform slider?

## 1.1 Basic Model: Error ~ Round

```{r}
#| label: analysis1-basic

# Linear mixed-effects model: error decreases over time
model1_basic <- lmer(
  error ~ idx + (1 + idx | gameID/playerID),
  data = d.all
)

cat("=== Model 1: Error ~ Round ===\n")
summary(model1_basic)

# Extract key coefficient
coef_round <- fixef(model1_basic)["idx"]
ci_round <- confint(model1_basic, parm = "idx", method = "Wald")
cat("\nRound effect: b =", round(coef_round, 3),
    ", 95% CI [", round(ci_round[1], 3), ",", round(ci_round[2], 3), "]\n")
```

## 1.2 Condition Effect: Error ~ Round × Condition

```{r}
#| label: analysis1-condition

# Model with condition
model1_condition <- lmer(
  error ~ idx * condition + (1 + idx | gameID/playerID),
  data = d.all
)

cat("=== Model 2: Error ~ Round × Condition ===\n")
summary(model1_condition)

# Model comparison
cat("\n=== Model Comparison ===\n")
anova(model1_basic, model1_condition)
```

## 1.3 Figure 5A: Error Over Time by Condition

```{r}
#| label: fig5a
#| fig-width: 8
#| fig-height: 5

# Compute bootstrapped means
error_by_round <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(condition, idx) %>%
  tidyboot_mean(error, na.rm = TRUE)

fig5a <- ggplot(error_by_round, aes(x = idx, y = empirical_stat, color = condition)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper, fill = condition),
              alpha = 0.2, color = NA) +
  geom_point(size = 2) +
  scale_color_manual(values = condition_colors,
                     labels = c("slider" = "Slider", "interactive" = "Chat")) +
  scale_fill_manual(values = condition_colors,
                    labels = c("slider" = "Slider", "interactive" = "Chat")) +
  labs(
    x = "Round",
    y = "Mean Absolute Error",
    color = "Condition",
    fill = "Condition",
    title = "A. Error decreases over time"
  ) +
  theme(legend.position = "bottom")

fig5a
```

## 1.4 Figure 5B: Final Round Error Distribution

```{r}
#| label: fig5b
#| fig-width: 6
#| fig-height: 5

# Final round only
d.final <- d.all %>%
  filter(idx == max(idx, na.rm = TRUE))

fig5b <- ggplot(d.final, aes(x = condition, y = error, fill = condition)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.2, alpha = 0.8, outlier.shape = NA) +
  scale_fill_manual(values = condition_colors,
                    labels = c("slider" = "Slider", "interactive" = "Chat")) +
  scale_x_discrete(labels = c("slider" = "Slider", "interactive" = "Chat")) +
  labs(
    x = "Condition",
    y = "Final Round Error",
    title = "B. Final round error distribution"
  ) +
  theme(legend.position = "none")

fig5b
```

---

# Analysis 2: Difficulty × Condition Interactions (Binned)

**Key question:** Does chat's advantage emerge specifically under epistemic difficulty?

## 2.1 Three-Way Hard Condition

```{r}
#| label: analysis2-hard-all

# Filter to all-hard (HHH) and all-easy (EEE) for clearest comparison
d.extreme <- d.all %>%
  filter(difficulty_profile %in% c("HHH", "EEE"))

cat("Sample sizes for extreme conditions:\n")
d.extreme %>%
  distinct(gameID, condition, difficulty_profile) %>%
  count(difficulty_profile, condition) %>%
  print()

# Model: Error ~ Condition × Difficulty Profile
model2_extreme <- lmer(
  error ~ condition * difficulty_profile + (1 | gameID/playerID),
  data = d.extreme
)

cat("\n=== Model: Extreme Difficulty Profiles ===\n")
summary(model2_extreme)
```

## 2.2 Full Difficulty Profile Analysis

```{r}
#| label: analysis2-full

# All 8 difficulty profiles
model2_full <- lmer(
  error ~ condition * difficulty_profile + (1 | gameID/playerID),
  data = d.all
)

cat("=== Model: All Difficulty Profiles ===\n")
summary(model2_full)

# Estimated marginal means
cat("\n=== Estimated Marginal Means ===\n")
emmeans(model2_full, ~ condition | difficulty_profile)
```

## 2.3 Two-Way Difficulty Combinations

```{r}
#| label: analysis2-twoway

# Create two-way difficulty indicators
d.all <- d.all %>%
  mutate(
    hard_entropy_rep = (!entropy_easy) & (!rep_easy),
    hard_critters_rep = (!critters_easy) & (!rep_easy),
    hard_entropy_critters = (!entropy_easy) & (!critters_easy)
  )

# Hard entropy + hard representativeness
cat("=== Hard Entropy × Hard Representativeness ===\n")
d.subset <- d.all %>% filter(hard_entropy_rep)
if(nrow(d.subset) > 100) {
  model_ent_rep <- lmer(error ~ condition + (1 | gameID/playerID), data = d.subset)
  print(summary(model_ent_rep))
}

# Hard critters + hard representativeness
cat("\n=== Hard Critters × Hard Representativeness ===\n")
d.subset <- d.all %>% filter(hard_critters_rep)
if(nrow(d.subset) > 100) {
  model_crit_rep <- lmer(error ~ condition + (1 | gameID/playerID), data = d.subset)
  print(summary(model_crit_rep))
}

# Hard entropy + hard critters (easy representativeness)
cat("\n=== Hard Entropy × Hard Critters (Easy Rep) ===\n")
d.subset <- d.all %>% filter(hard_entropy_critters & rep_easy)
if(nrow(d.subset) > 100) {
  model_ent_crit <- lmer(error ~ condition + (1 | gameID/playerID), data = d.subset)
  print(summary(model_ent_crit))
}
```

## 2.4 Figure 6A: Difficulty Profile × Condition

```{r}
#| label: fig6a
#| fig-width: 10
#| fig-height: 6

# Compute mean error by difficulty profile and condition (bootstrapped CIs)
difficulty_summary <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(difficulty_profile, condition) %>%
  tidyboot_mean(error, na.rm = TRUE) %>%
  # Order by number of hard dimensions
  mutate(
    n_hard = str_count(difficulty_profile, "H"),
    difficulty_profile = fct_reorder(difficulty_profile, n_hard)
  )

fig6a <- ggplot(difficulty_summary,
                aes(x = difficulty_profile, y = empirical_stat, fill = condition)) +
  geom_col(position = position_dodge(0.8), width = 0.7) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                position = position_dodge(0.8), width = 0.2) +
  scale_fill_manual(values = condition_colors,
                    labels = c("slider" = "Slider", "interactive" = "Chat")) +
  labs(
    x = "Difficulty Profile (Critters/Entropy/Representativeness)",
    y = "Mean Absolute Error",
    fill = "Condition",
    title = "A. Error by difficulty profile and condition"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

fig6a
```

---

# Analysis 3: Continuous Difficulty Predictors

**Key question:** Which difficulty dimension most strongly moderates chat's advantage?

## 3.1 Min-Max Scale Predictors

Using the **actual network metrics** for scaling:

```{r}
#| label: analysis3-scale

# Min-max scale continuous predictors (using actual network metrics)
d.all <- d.all %>%
  mutate(
    critters_scaled = (nCrittersGame_actual - min(nCrittersGame_actual, na.rm = TRUE)) /
                      (max(nCrittersGame_actual, na.rm = TRUE) - min(nCrittersGame_actual, na.rm = TRUE)),
    entropy_scaled = (entropy_actual - min(entropy_actual, na.rm = TRUE)) /
                     (max(entropy_actual, na.rm = TRUE) - min(entropy_actual, na.rm = TRUE)),
    rep_scaled = (representation_actual - min(representation_actual, na.rm = TRUE)) /
                 (max(representation_actual, na.rm = TRUE) - min(representation_actual, na.rm = TRUE))
  )

cat("Scaled predictor ranges:\n")
cat("Critters:", range(d.all$critters_scaled, na.rm = TRUE), "\n")
cat("Entropy:", range(d.all$entropy_scaled, na.rm = TRUE), "\n")
cat("Representativeness:", range(d.all$rep_scaled, na.rm = TRUE), "\n")
```

## 3.2 Full Interaction Model

```{r}
#| label: analysis3-full

# Full model with all interactions
model3_full <- lmer(
  error ~ condition * critters_scaled * entropy_scaled * rep_scaled +
    (1 | gameID),
  data = d.all
)

cat("=== Full Interaction Model ===\n")
summary(model3_full)
```

## 3.3 Extract Key Interactions

```{r}
#| label: analysis3-interactions

# Simpler model focusing on two-way interactions with condition
model3_simple <- lmer(
  error ~ condition * critters_scaled +
          condition * entropy_scaled +
          condition * rep_scaled +
    (1 | gameID/playerID),
  data = d.all
)

cat("=== Two-Way Interactions with Condition ===\n")
summary(model3_simple)

# Extract coefficients for figure
coef_table <- tidy(model3_simple, effects = "fixed", conf.int = TRUE) %>%
  filter(str_detect(term, ":")) %>%
  mutate(
    dimension = case_when(
      str_detect(term, "critters") ~ "Sample Size",
      str_detect(term, "entropy") ~ "Entropy",
      str_detect(term, "rep") ~ "Representativeness"
    )
  )

print(coef_table)
```

## 3.4 Figure 6B: Interaction Coefficients

```{r}
#| label: fig6b
#| fig-width: 6
#| fig-height: 4

# Coefficient plot
fig6b <- ggplot(coef_table, aes(x = reorder(dimension, abs(estimate)),
                                 y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    x = "",
    y = "Condition × Dimension Interaction (β)",
    title = "B. Interaction coefficients"
  )

fig6b
```

---

# Analysis 4: Message Content

**Key question:** What aspects of chat communication predict accuracy?

## 4.1 Load Coded Messages

```{r}
#| label: load-messages
#| message: false

# Load coded message data
d.messages <- read_csv(here('analysis/exp2/message_data/coded.csv'))

cat("=== MESSAGE DATA ===\n")
cat("Total messages:", nrow(d.messages), "\n")
cat("Unique games:", n_distinct(d.messages$gameID), "\n")

# Coding categories
cat("\nCoding categories:\n")
d.messages %>%
  count(coding) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  print()
```

## 4.2 Aggregate Message Counts per Game

```{r}
#| label: message-counts

# Count message types per game
message_counts <- d.messages %>%
  group_by(gameID) %>%
  summarise(
    total_messages = n(),
    pn_count = sum(coding == "pn", na.rm = TRUE),  # personal numeric
    pb_count = sum(coding == "pb", na.rm = TRUE),  # personal belief
    sn_count = sum(coding == "sn", na.rm = TRUE),  # social numeric
    sb_count = sum(coding == "sb", na.rm = TRUE),  # social belief
    o_count = sum(coding == "o", na.rm = TRUE),    # other
    q_count = sum(coding == "q", na.rm = TRUE),    # question
    .groups = "drop"
  ) %>%
  mutate(
    # Proportions
    p_social = (sn_count + sb_count) / total_messages,
    p_personal = (pn_count + pb_count) / total_messages
  )

cat("Message count summary:\n")
summary(message_counts %>% select(total_messages, pn_count:q_count))
```

## 4.3 Merge with Game-Level Error

```{r}
#| label: merge-messages

# Compute final-round error per game
game_error <- d.all %>%
  filter(idx == max(idx, na.rm = TRUE)) %>%
  group_by(gameID) %>%
  summarise(
    avg_error = mean(error, na.rm = TRUE),
    .groups = "drop"
  )

# Merge
d.message_analysis <- message_counts %>%
  inner_join(game_error, by = "gameID")

cat("Games with both message and error data:", nrow(d.message_analysis), "\n")
```

## 4.4 Social Information Predicts Accuracy

```{r}
#| label: analysis4-social

# Model: Final error ~ proportion social information
model4_social <- lm(
  log(avg_error + 1) ~ p_social + p_personal,
  data = d.message_analysis
)

cat("=== Model: Social Information → Error ===\n")
summary(model4_social)

# By message type
model4_types <- lm(
  log(avg_error + 1) ~ pn_count + pb_count + sn_count + sb_count,
  data = d.message_analysis
)

cat("\n=== Model: Message Types → Error ===\n")
summary(model4_types)
```

## 4.5 Figure 7B: Message Type Frequencies

```{r}
#| label: fig7b
#| fig-width: 6
#| fig-height: 4

# Frequency plot - filter to main coding categories only
freq_data <- d.messages %>%
  filter(coding %in% c("pn", "pb", "sn", "sb", "o", "q")) %>%
  count(coding) %>%
  mutate(
    label = case_when(
      coding == "pn" ~ "Personal\nNumeric",
      coding == "pb" ~ "Personal\nBelief",
      coding == "sn" ~ "Social\nNumeric",
      coding == "sb" ~ "Social\nBelief",
      coding == "o" ~ "Other",
      coding == "q" ~ "Question",
      TRUE ~ coding
    ),
    type = ifelse(coding %in% c("sn", "sb"), "Social", "Personal/Other")
  )

fig7b <- ggplot(freq_data, aes(x = reorder(label, -n), y = n, fill = type)) +
  geom_col() +
  scale_fill_manual(values = c("Social" = "#0072B2", "Personal/Other" = "gray60")) +
  labs(
    x = "",
    y = "Count",
    fill = "",
    title = "B. Message type frequencies"
  ) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 9))

fig7b
```

## 4.6 Figure 7C: Social Information vs Error

```{r}
#| label: fig7c
#| fig-width: 6
#| fig-height: 5

fig7c <- ggplot(d.message_analysis, aes(x = p_social, y = avg_error)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "#0072B2") +
  labs(
    x = "Proportion Social Information (sn + sb)",
    y = "Final Round Error",
    title = "C. Social information predicts accuracy"
  )

fig7c
```

---

# Combine and Save Figures

## Figure 5: Combined Results Figure (Time + Difficulty)

This is the main "money plot" combining error over time with the three difficulty dimensions.

```{r}
#| label: save-fig5
#| fig-width: 12
#| fig-height: 3.5

# Panel A: Error over time (bootstrapped CIs)
error_by_round <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(condition, idx) %>%
  tidyboot_mean(error, na.rm = TRUE)

fig5_a <- ggplot(error_by_round, aes(x = idx, y = empirical_stat, color = condition)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper, fill = condition),
              alpha = 0.2, color = NA) +
  geom_point(size = 2) +
  scale_color_manual(values = condition_colors,
                     labels = c("interactive" = "Chat", "slider" = "Slider"),
                     name = "Condition") +
  scale_fill_manual(values = condition_colors,
                    labels = c("interactive" = "Chat", "slider" = "Slider"),
                    name = "Condition") +
  labs(x = "Round", y = "Mean Absolute Error", tag = "A") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(24, 40))

# Panel B: Sample Size (bootstrapped CIs)
summary_critters <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(critters_bin, condition) %>%
  tidyboot_mean(error, na.rm = TRUE)

fig5_b <- ggplot(summary_critters, aes(x = critters_bin, y = empirical_stat, color = condition, group = condition)) +
  geom_line(linewidth = 1) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.1) +
  scale_color_manual(values = condition_colors) +
  scale_x_discrete(labels = c("Few" = "Few\n▪▪", "Many" = "Many\n▪▪▪▪▪")) +
  labs(x = "Sample Size", y = "", tag = "B") +
  theme(legend.position = "none", axis.text.x = element_text(size = 9, lineheight = 0.9)) +
  coord_cartesian(ylim = c(24, 36))

# Panel C: Entropy (bootstrapped CIs)
summary_entropy <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(entropy_bin, condition) %>%
  tidyboot_mean(error, na.rm = TRUE)

fig5_c <- ggplot(summary_entropy, aes(x = entropy_bin, y = empirical_stat, color = condition, group = condition)) +
  geom_line(linewidth = 1) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.1) +
  scale_color_manual(values = condition_colors) +
  scale_x_discrete(labels = c("Low" = "Low\n●○○○", "High" = "High\n●●●●")) +
  labs(x = "Distribution", y = "", tag = "C") +
  theme(legend.position = "none", axis.text.x = element_text(size = 9, lineheight = 0.9)) +
  coord_cartesian(ylim = c(24, 36))

# Panel D: Representativeness (bootstrapped CIs)
summary_rep <- d.all %>%
  filter(!is.na(error)) %>%
  group_by(rep_bin, condition) %>%
  tidyboot_mean(error, na.rm = TRUE)

fig5_d <- ggplot(summary_rep, aes(x = rep_bin, y = empirical_stat, color = condition, group = condition)) +
  geom_line(linewidth = 1) + geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.1) +
  scale_color_manual(values = condition_colors,
                     labels = c("interactive" = "Chat", "slider" = "Slider"),
                     name = "Condition") +
  scale_x_discrete(labels = c("Representative" = "Representative\n◐◐◐◐", "Unrepresentative" = "Unrepresentative\n●○●○")) +
  labs(x = "Representativeness", y = "", tag = "D") +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 8, lineheight = 0.9)) +
  coord_cartesian(ylim = c(24, 36))

# Combine
fig5 <- (fig5_a | fig5_b | fig5_c | fig5_d) +
  plot_layout(widths = c(1.2, 1, 1, 1.1), guides = "collect") &
  theme(legend.position = "bottom")

fig5

ggsave(here("figures/fig5.pdf"), fig5, width = 12, height = 3.5)
ggsave(here("figures/fig5.png"), fig5, width = 12, height = 3.5, dpi = 300)
cat("Saved Figure 5 (combined results)\n")
```

## Figure 6: Message Content

```{r}
#| label: save-fig6
#| fig-width: 10
#| fig-height: 4

# Panel A: Message type frequencies
freq_data <- d.messages %>%
  filter(coding %in% c("pn", "pb", "sn", "sb", "o", "q")) %>%
  count(coding) %>%
  mutate(
    label = case_when(
      coding == "pn" ~ "Personal\nNumeric",
      coding == "pb" ~ "Personal\nBelief",
      coding == "sn" ~ "Social\nNumeric",
      coding == "sb" ~ "Social\nBelief",
      coding == "o" ~ "Other",
      coding == "q" ~ "Question"
    ),
    type = ifelse(coding %in% c("sn", "sb"), "Social", "Personal/Other")
  )

fig6_a <- ggplot(freq_data, aes(x = reorder(label, -n), y = n, fill = type)) +
  geom_col() +
  scale_fill_manual(values = c("Social" = "#0072B2", "Personal/Other" = "gray60")) +
  labs(x = "", y = "Count", fill = "", tag = "A") +
  theme(legend.position = "bottom", axis.text.x = element_text(size = 9))

# Panel B: Social information vs error
fig6_b <- ggplot(d.message_analysis, aes(x = p_social, y = avg_error)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "#0072B2") +
  labs(x = "Proportion Social Information", y = "Final Round Error", tag = "B")

fig6 <- fig6_a + fig6_b + plot_layout(widths = c(1, 1))

fig6

ggsave(here("figures/fig6.pdf"), fig6, width = 10, height = 4)
ggsave(here("figures/fig6.png"), fig6, width = 10, height = 4, dpi = 300)
cat("Saved Figure 6 (message content)\n")
```

---

# Summary Statistics for Manuscript

```{r}
#| label: summary-stats

cat("\n========================================\n")
cat("SUMMARY STATISTICS FOR MANUSCRIPT\n")
cat("========================================\n\n")

cat("=== Sample ===\n")
cat("N participants:", n_distinct(d.all$playerID), "\n")
cat("N games:", n_distinct(d.all$gameID), "\n")

cat("\n=== Analysis 1: Error over time ===\n")
cat("Round effect: b =", round(fixef(model1_basic)["idx"], 2), "\n")

cat("\n=== Analysis 2: Condition effect ===\n")
# Extract from model1_condition

cat("\n=== Analysis 3: Interactions ===\n")
print(coef_table %>% select(dimension, estimate, conf.low, conf.high, p.value))

cat("\n=== Analysis 4: Message content ===\n")
cat("Social information effect from model4_social\n")
```
